InnoDB's Full Text Search Dynamic article with Shiny/R

### 2016-01-04 article.rmd
Some cosmetic changes before push.

---
title: "InnoDB Full text Search hacking with Shiny/R"
author: "Emanuel Calvo - Pythian"
date: "December 22, 2015"
output: html_document
runtime: shiny
---

@@

- Use QUERY EXPANSION only if you are interested to search relations over exact matches. Remember that the field
  size is crucial when using this.
- FTS is not the best fit for exact string matches in single columns. You don't want to use FTS for searching emails in
  a single column, name and lastname fields , i.e. For those, you'll probably use other techniques as reverse searches or
  exact match operator (=).
- Keep your FTS indexes short. Do not add ALL the text columns. Parse first from your application the user search and 
  adapt the query. 
- If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the
  FTS functions to avoid double executions. You can do this using something like:
  `match(content,title) against ("first (<second >third)") > 1 `
  Generally, scores lower than 1 can be ignored when using boolean or natural mode searches.
- `OPTIMIZE TABLE` does a rebuild of the table. To avoid this, set `innodb_optimize_fulltext_only=1` in order to do an incremental
  maintance on the table.
- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE.
- If you plan to order by rank, it is not necessary to specify the clause `ORDER BY` as MySQL does the order forcelly internally. Also,
  the behavior is different from the default as it returns the heaviest at the top (like ORDER BY rank DESC).
- If you come from MyISAM's FTS implementation, recall that the ranking scoring is different.
- Create the FULLTEXT index after the data is loaded [InnoDB bulk load](http://dev.mysql.com/doc/refman/5.7/en/optimizing-innodb-bulk-data-loading.html). When restoring FTS backups, you will probably hit the "ERROR 182 (HY000) at line nn: Invalid InnoDB FTS Doc ID".
- Try to don't use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
  it consumes a considerably amount of CPU. MySQL orders by the latest expression in the WHERE clause. [WL#7123](https://dev.mysql.com/worklog/task/?id=7123)
- Also, if avoiding the rank information in the projection (SELECT clause) and using other aggregations like `count(*)`, will use
  the "no ranking" FT_hints. The `limit` hint won't be used if invoked explicitely an `ORDER BY` and th `MATCH` clause in the projection.

@@

                DataResults = reactive({
                   f<-function(x) {as.character(input$filtering)}
                  sql <- paste("SELECT FTS_DOC_ID, bookid,
                              MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,"), content
                              FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") LIMIT 50", sep = "")

@@

                      sql <- paste("SELECT sourceTable,
                                      max(scoreBoolean) as maxBoolean,
                                      max(scoreQE) as maxQE,
                                      -- max(scoreNL) as maxNL,
                                      avg(scoreBoolean) as avgBoolean,
                                      avg(scoreQE) as avgQE,
                                      -- avg(scoreNL) as avgNL,
                                      avg(lengthContent) avgFieldLength,
                                      -- avg(scoreQE) / count(sourceTable) as weight,
                                      count(sourceTable) as countOccurrences
                              FROM
                              (
                              SELECT \"byParagraph\" as sourceTable,
                                    FTS_DOC_ID,
                                    match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
                                    match(BC.content)
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
                                    match(BC.content)
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE) scoreNL,
                                     length(content) as lengthContent
                              FROM    bookContent BC
                              WHERE
                                match(BC.content) against (\"" ,f(x)," \" ",input$ftsMode,")
                              UNION
                              SELECT \"byLine\" as sourceTable,
                                      FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
                                     match(BL.content)
                                      against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
                                    match(BL.content)
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE) scoreNL,
                                     length(content) as lengthContent
                              FROM
                                      bookContentByLine BL
                              WHERE
                                match(BL.content) against (\"", f(x) ,"\" ",input$ftsMode,")
                              ) unionTable
                              GROUP BY sourceTable
                              ;" , sep = "")

>>>>

---
title: "InnoDB's Full Text Search Dynamic article with Shiny/R"
author: "Emanuel Calvo - Pythian"
date: "January 3rd, 2016"
output: html_document
runtime: shiny
---

@@

- Use QUERY EXPANSION only if you are interested to search relations over exact matches. Remember that the field
  size is crucial when using this.
- FTS is not the best fit for exact string matches in single columns. You don't want to use FTS for searching emails in
  a single column, name and lastname fields , i.e. For those, you'll probably use other techniques as reverse searches or
  exact match operator (=).
- Keep your FTS indexes short. Do not add ALL the text columns. Parse first from your application the user search and 
  adapt the query. 
- If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the
  FTS functions to avoid double executions. You can do this using something like:
  `match(content,title) against ("first (<second >third)") > 1 `
  Generally, scores lower than 1 can be ignored when using boolean or natural mode searches.
- `OPTIMIZE TABLE` does a rebuild of the table. To avoid this, set `innodb_optimize_fulltext_only=1` in order to do an incremental
  maintance on the table.
- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE. This affects the ranking score (try __+bad (<feeling >thing)__ i.e.)
- If you plan to order by rank, it is not necessary to specify the clause `ORDER BY` as InnoDB does the order after retrieve the doc ids . Also,the behavior is different from the default as it returns the heaviest at the top (like an __ORDER BY rank DESC__).
- If you come from MyISAM's FTS implementation, recall that the ranking scoring is different.
- Create the FULLTEXT index after the data is loaded [InnoDB bulk load](http://dev.mysql.com/doc/refman/5.7/en/optimizing-innodb-bulk-data-loading.html). When restoring FTS backups, you will probably hit the "ERROR 182 (HY000) at line nn: Invalid InnoDB FTS Doc ID".
- Try to avoid using use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
  it consumes a considerably amount of CPU. InnoDB orders by the latest expression in the WHERE clause. [WL#7123](https://dev.mysql.com/worklog/task/?id=7123)
- Also, if avoiding the rank information in the projection (SELECT clause) and using other aggregations like `count(*)`, will use
  the "no ranking" FT_hints. The `limit` hint won't be used if invoked explicitely an `ORDER BY` and th `MATCH` clause in the projection.

@@

                DataResults = reactive({
                   f<-function(x) {as.character(input$filtering)}
                  sql <- paste("SELECT FTS_DOC_ID, bookid,
                              MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") as rankScore, content
                              FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") LIMIT 50", sep = "")

@@

                      sql <- paste("SELECT sourceTable,
                                      max(scoreBoolean) as maxBoolean,
                                      max(scoreQE) as maxQE,
                                      max(scoreNL) as maxNL,
                                      avg(scoreBoolean) as avgBoolean,
                                      avg(scoreQE) as avgQE,
                                      avg(scoreNL) as avgNL,
                                      avg(lengthContent) avgFieldLength,
                                      -- avg(scoreQE) / count(sourceTable) as weight,
                                      count(sourceTable) as countOccurrences
                              FROM
                              (
                              SELECT \"byParagraph\" as sourceTable,
                                    FTS_DOC_ID,
                                    match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
                                    match(BC.content)
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
                                    match(BC.content)
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE) scoreNL,
                                     length(content) as lengthContent
                              FROM    bookContent BC
                              WHERE
                                match(BC.content) against (\"" ,f(x)," \" ",input$ftsMode,")
                              UNION
                              SELECT \"byLine\" as sourceTable,
                                      FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
                                     match(BL.content)
                                      against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
                                    match(BL.content) 
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE) scoreNL,
                                     length(content) as lengthContent
                              FROM
                                      bookContentByLine BL
                              WHERE
                                match(BL.content) against (\"", f(x) ,"\" ",input$ftsMode,")
                              ) unionTable
                              GROUP BY sourceTable
                              ;" , sep = "")
